---
title: "EDA_UDS"
author: "Aubree Krager"
date: "2024-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup,include=FALSE,results=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hold')
#Load Libraries and functions Dr. Abdoli made
library(magrittr)
library(dplyr)
library(tidyr)
library(tidyselect)
library(kableExtra)
library(NACCdata)
library(glmnet)
library(car)
library(tibble)
library(ggmosaic)
library(ROCit)
library(pROC)
library(caret)
library()

colTrim <- function(x, width){
  x <- trimws(x)
  sapply(x, strtrim, width=width)
}
source('describeDS.R')
source('briefSummary.R')
source('xval.R')
```

### Get Data:  
These are the variables that will be used in the model. The order of variables are chosen for convenience of comparing coefficients; the order should not significantly affect the overall results.

```{r}
col2clean <- c('NACCID', 'NACCALZD', 'NACCAGE', 'BPDIAS', 'BPSYS',
               'HXHYPER','HYPERCHO', 'HXSTROKE','CVHATT', 'CVCHF',
               'CVAFIB','NACCBMI', 'DIABETES', 'B12DEF', 'DEP2YRS', 
               'SMOKYRS', 'TOBAC100', 'PACKSPER', 'ALCOHOL', 
               'NACCNIHR', 'HISPANIC','EDUC', 'SEX', 'MARISTAT',
               'NACCVNUM', 'NACCNIHR_updated')
cat(sprintf('{%s}', paste(col2clean, collapse = ', ')), sep='\n')
```

### Variable Specification
The variable cleaning specifications are shown below:

```{r}
#reading the cleaning instructions
glmDS0 <- read.csv('cleanUDS_EDA.csv')

#making table
glmDS0 %>% dplyr::select(-c(Rename, Comment, MissingValComment)) %>% 
  dplyr::filter(GLM!="", Name %in% col2clean) %>% 
  mutate(CurrentVal=colTrim(CurrentVal, 20)) %>% 
  mutate(NewLevel=colTrim(NewLevel, 20)) %>% 
  kbl(caption = 'UDS Variables for the Study') %>% 
  kable_styling(full_width = FALSE,
                bootstrap_options = c("striped", "hover", "condensed",
                                      "responsive")) %>% 
  column_spec(2, width_max = '2in') %>% 
  column_spec(3, width_max = '2in') %>% 
  scroll_box(height = "400px")
```

# Create the data frame
```{r}
#make data frame
X <- NACCdata::cleanDS(
  cleanDF = glmDS0 %>% filter(GLM!="", Name %in% col2clean)
    ) 

#checking race statements
#how many people are multiracial and Hispanic
nrow(X %>% 
  filter(NACCNIHR == "Multiracial" & HISPANIC == "Hispanic"))
#how many people are any race and hispanic? 
nrow(X %>% 
  filter(NACCNIHR != "" & HISPANIC == "Hispanic"))
#how many people have a NA race value and are non hispanic
nrow(X %>% 
  filter(is.na(NACCNIHR) & HISPANIC == "non-Hispanic"))
#how many people  have a NA race value and are hispanic
nrow(X %>% 
  filter(is.na(NACCNIHR) & HISPANIC == "Hispanic"))

#make updated race variable
X <- X %>% 
  mutate(NACCNIHR_updated = case_when(
    NACCNIHR != "" & HISPANIC == "Hispanic" ~ "Multiracial",
    is.na(NACCNIHR) & HISPANIC == "non-Hispanic" ~ "Other",
    is.na(NACCNIHR) & HISPANIC == "Hispanic" ~ "Hispanic",
    TRUE ~ NACCNIHR
  ))

#checking to see if correct
table(X$NACCNIHR_updated)


#set reference category
X$NACCNIHR_updated <- relevel(factor(X$NACCNIHR_updated), ref = "White")

#how many records
X <- X %>% filter(NACCVNUM==1)
cat(sprintf('%d records is read.\n', dim(X)[1]))
```

## Incomplete Cases {.tabset}
The incomplete cases are dropped and the counts of NA's before and after are reported. Then, variables are summarized in a brief format more appropriate for publications. 

### Counting NAs
```{r}
cat('Before removing incomplete observations:\n\n')
(naCounts <- sapply(colnames(X), \(v) sum(is.na(X[[v]]))))

X <- X %>% 
  drop_na(ALCOHOL) %>% 
  drop_na(EDUC) %>% 
  drop_na(MARISTAT) %>% 
  drop_na(NACCBMI) %>% 
  drop_na(PACKSPER) %>% 
  drop_na(B12DEF) %>% 
  drop_na(BPDIAS) %>% 
  drop_na(BPSYS) %>% 
  drop_na(CVAFIB) %>% 
  drop_na(CVCHF) %>% 
  drop_na(CVHATT) %>% 
  drop_na(DEP2YRS) %>% 
  drop_na(DIABETES) %>% 
  drop_na(HXHYPER) %>% 
  drop_na(HXSTROKE) %>% 
  drop_na(HYPERCHO) %>% 
  drop_na(NACCNIHR_updated)

cat('\n\n\nAfter removing incomplete observations:\n\n')
sapply(colnames(X), \(v) sum(is.na(X[[v]])))

```

## Only AD and Normal Cognition
```{r}
X18 <- X %>% mutate(NACCALZD=factor(NACCALZD, c("Normal", "Alzheimer"))) %>%
  filter(!is.na(NACCALZD))

nrow(X18)
```

## Single Numeric Variables

# Boxplot of Age by Diagnosis
```{r}
  ggplot(data = X18, aes(x = NACCALZD, y = NACCAGE)) +
  geom_boxplot(fill = "#009E73") +
  labs(title = "Age by Diagnosis", 
       y = "Age (yr)",
       x = "Diagnosis") +
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Boxplot of Years of Education by Diagnosis
```{r}
  ggplot(data = X18, aes(x = NACCALZD, y = EDUC)) +
  geom_boxplot(fill = "#009E73") +
  labs(title = "Years of Education by Diagnosis", 
       y = "Years of Education",
       x = "Diagnosis") +
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Boxplot of BMI by Diagnosis
```{r}
  ggplot(data = X18, aes(x = NACCALZD, y = NACCBMI)) +
  geom_boxplot(fill = "#009E73") +
  labs(title = "BMI by Diagnosis", 
       y = "Body Mass Index (BMI)",
       x = "Diagnosis") +
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Boxplot of Diastolic Blood Pressure by Diagnosis
```{r}
  ggplot(data = X18, aes(x = NACCALZD, y = BPDIAS)) +
  geom_boxplot(fill = "#009E73") +
  labs(title = "Diastolic Blood Pressure by Diagnosis", 
       y = "Diastolic Blood Pressure",
       x = "Diagnosis") +
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Boxplot of Systolic Blood Pressure by Diagnosis
```{r}
  ggplot(data = X18, aes(x = NACCALZD, y = BPSYS)) +
  geom_boxplot(fill = "#009E73") +
  labs(title = "Systolic Blood Pressure by Diagnosis", 
       y = "Systolic Blood Pressure",
       x = "Diagnosis") +
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

## Two Numerical Variables (Age vs. Variable)

Note, these EDA groaphs are for a patients first visit and only considers patients with normal cognition and AD. 

# Scatter plot of Age and Years of Education by Diagnosis
```{r}
  ggplot(data = X18, aes(x = NACCAGE, y = EDUC, color = NACCALZD)) +
  geom_point() +
  labs(title = "Age and Years of Education by Diagnosis", 
       y = "Years of Education",
       x = "Age (yr)", color = "Diagnosis") +
  geom_smooth(method = "lm", se = FALSE) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))+
  scale_color_manual(values = c("#edcb69", "#385a93"))
```

# Scatter plot of Age and BMI by Diagnosis
```{r}
  ggplot(data = X18, aes(x = NACCAGE, y = NACCBMI, color = NACCALZD)) +
  geom_point() +
  labs(title = "Age and BMI by Diagnosis", 
       y = "Body Mass Index (BMI)",
       x = "Age (yr)", color = "Diagnosis") +
  geom_smooth(method = "lm", se = FALSE) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))+
  scale_color_manual(values = c("#edcb69", "#385a93"))
```

# Scatter plot of Age and Diastolic Blood Pressure by Diagnosis
```{r}
  ggplot(data = X18, aes(x = NACCAGE, y = BPDIAS, color = NACCALZD)) +
  geom_point() +
  labs(title = "Age and Diastolic Blood Pressure by Diagnosis", 
       y = "Diastolic Blood Pressure",
       x = "Age (yr)", color = "Diagnosis") +
  geom_smooth(method = "lm", se = FALSE) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))+
  scale_color_manual(values = c("#edcb69", "#385a93"))
```

# Scatter plot of Age and Systolic Blood Pressure by Diagnosis
```{r}
  ggplot(data = X18, aes(x = NACCAGE, y = BPSYS, color = NACCALZD)) +
  geom_point() +
  labs(title = "Age and Systolic Blood Pressure by Diagnosis", 
       y = "Systolic Blood Pressure",
       x = "Age (yr)", color = "Diagnosis") +
  geom_smooth(method = "lm", se = FALSE) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))+
  scale_color_manual(values = c("#edcb69", "#385a93"))
```

# Categorical Variables

# Mosaic Plot of Alcohol Abuse and Diagnosis
```{r}
# Calculate the proportion table
alcohol.prop <- prop.table(table(X18$NACCALZD, X18$ALCOHOL), margin = 1)

# Convert the proportion table to a data frame
alcohol.prop_df <- as.data.frame(alcohol.prop)

# Rename the columns for clarity
colnames(alcohol.prop_df) <- c("Diagnosis", "Alcohol", "Proportion")

print(alcohol.prop_df)

ggplot(alcohol.prop_df, aes(x = Alcohol, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Alcohol") +
  ylab("Diagnosis") +
  ggtitle("Alcohol Abuse by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of Marital Status and Diagnosis

```{r}
# Calculate the proportion table
MARISTAT.prop <- prop.table(table(X18$NACCALZD, X18$MARISTAT), margin = 1)

# Convert the proportion table to a data frame
MARISTAT.prop_df <- as.data.frame(MARISTAT.prop)

# Rename the columns for clarity
colnames(MARISTAT.prop_df) <- c("Diagnosis", "MARISTAT", "Proportion")

print(MARISTAT.prop_df)

ggplot(MARISTAT.prop_df, aes(x = MARISTAT, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Marital Status") +
  ylab("Diagnosis") +
  ggtitle("Marital Status by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of Average Number of Cigarette Packs Smoked per Day and Diagnosis

```{r}
# Calculate the proportion table
PACKSPER.prop <- prop.table(table(X18$NACCALZD, X18$PACKSPER), margin = 1)

# Convert the proportion table to a data frame
PACKSPER.prop_df <- as.data.frame(PACKSPER.prop)

# Rename the columns for clarity
colnames(PACKSPER.prop_df) <- c("Diagnosis", "PACKSPER", "Proportion")

print(PACKSPER.prop_df)

ggplot(PACKSPER.prop_df, aes(x = PACKSPER, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab(" Average Number of Cigarette Packs Smoked per Day") +
  ylab("Diagnosis") +
  ggtitle("Average Number of Cigarette Packs Smoked per Day by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of B12 Deficency and Diagnosis
```{r}
# Calculate the proportion table
B12DEF.prop <- prop.table(table(X18$NACCALZD, X18$B12DEF), margin = 1)

# Convert the proportion table to a data frame
B12DEF.prop_df <- as.data.frame(B12DEF.prop)

# Rename the columns for clarity
colnames(B12DEF.prop_df) <- c("Diagnosis", "B12DEF", "Proportion")

print(B12DEF.prop_df)

ggplot(B12DEF.prop_df, aes(x = B12DEF, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("B12 Deficency") +
  ylab("Diagnosis") +
  ggtitle("B12 Deficency by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of Atrial Fibrillation and Diagnosis
```{r}
# Calculate the proportion table
CVAFIB.prop <- prop.table(table(X18$NACCALZD, X18$CVAFIB), margin = 1)

# Convert the proportion table to a data frame
CVAFIB.prop_df <- as.data.frame(CVAFIB.prop)

# Rename the columns for clarity
colnames(CVAFIB.prop_df) <- c("Diagnosis", "CVAFIB", "Proportion")

print(CVAFIB.prop_df)

ggplot(CVAFIB.prop_df, aes(x = CVAFIB, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Atrial Fibrillation") +
  ylab("Diagnosis") +
  ggtitle("Atrial Fibrillation by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of Congesive Heart Failure and Diagnosis
```{r}
# Calculate the proportion table
CVCHF.prop <- prop.table(table(X18$NACCALZD, X18$CVCHF), margin = 1)

# Convert the proportion table to a data frame
CVCHF.prop_df <- as.data.frame(CVCHF.prop)

# Rename the columns for clarity
colnames(CVCHF.prop_df) <- c("Diagnosis", "CVCHF", "Proportion")

print(CVCHF.prop_df)

ggplot(CVCHF.prop_df, aes(x = CVCHF, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Congesive Heart Failure") +
  ylab("Diagnosis") +
  ggtitle("Congesive Heart Failure by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```
  
# Mosaic Plot of Heart Attack/Cardiac Arrest and Diagnosis
```{r}
# Calculate the proportion table
CVHATT.prop <- prop.table(table(X18$NACCALZD, X18$CVHATT), margin = 1)

# Convert the proportion table to a data frame
CVHATT.prop_df <- as.data.frame(CVHATT.prop)

# Rename the columns for clarity
colnames(CVHATT.prop_df) <- c("Diagnosis", "CVHATT", "Proportion")

print(CVHATT.prop_df)

ggplot(CVHATT.prop_df, aes(x = CVHATT, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Heart Attack/Cardiac Arrest") +
  ylab("Diagnosis") +
  ggtitle("Heart Attack/Cardiac Arrest by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of Depression and Diagnosis
```{r}
# Calculate the proportion table
DEP2YRS.prop <- prop.table(table(X18$NACCALZD, X18$DEP2YRS), margin = 1)

# Convert the proportion table to a data frame
DEP2YRS.prop_df <- as.data.frame(DEP2YRS.prop)

# Rename the columns for clarity
colnames(DEP2YRS.prop_df) <- c("Diagnosis", "DEP2YRS", "Proportion")

print(DEP2YRS.prop_df)

ggplot(DEP2YRS.prop_df, aes(x = DEP2YRS, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Depression") +
  ylab("Diagnosis") +
  ggtitle("Depression by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```
  
# Mosaic Plot of Diabetes and Diagnosis
```{r}
# Calculate the proportion table
DIABETES.prop <- prop.table(table(X18$NACCALZD, X18$DIABETES), margin = 1)

# Convert the proportion table to a data frame
DIABETES.prop_df <- as.data.frame(DIABETES.prop)

# Rename the columns for clarity
colnames(DIABETES.prop_df) <- c("Diagnosis", "DIABETES", "Proportion")

print(DIABETES.prop_df)

ggplot(DIABETES.prop_df, aes(x = DIABETES, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Diabetes") +
  ylab("Diagnosis") +
  ggtitle("Diabetes by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```
  
# Mosaic Plot of Hypertension and Diagnosis
```{r}
# Calculate the proportion table
HXHYPER.prop <- prop.table(table(X18$NACCALZD, X18$HXHYPER), margin = 1)

# Convert the proportion table to a data frame
HXHYPER.prop_df <- as.data.frame(HXHYPER.prop)

# Rename the columns for clarity
colnames(HXHYPER.prop_df) <- c("Diagnosis", "HXHYPER", "Proportion")

print(HXHYPER.prop_df)

ggplot(HXHYPER.prop_df, aes(x = HXHYPER, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Hypertension") +
  ylab("Diagnosis") +
  ggtitle("Hypertension by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of Stroke and Diagnosis
```{r}
# Calculate the proportion table
HXSTROKE.prop <- prop.table(table(X18$NACCALZD, X18$HXSTROKE), margin = 1)

# Convert the proportion table to a data frame
HXSTROKE.prop_df <- as.data.frame(HXSTROKE.prop)

# Rename the columns for clarity
colnames(HXSTROKE.prop_df) <- c("Diagnosis", "HXSTROKE", "Proportion")

print(HXSTROKE.prop_df)

ggplot(HXSTROKE.prop_df, aes(x = HXSTROKE, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Stroke") +
  ylab("Diagnosis") +
  ggtitle("Stroke by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of Hypercholesterolemia and Diagnosis
```{r}
# Calculate the proportion table
HYPERCHO.prop <- prop.table(table(X18$NACCALZD, X18$HYPERCHO), margin = 1)

# Convert the proportion table to a data frame
HYPERCHO.prop_df <- as.data.frame(HYPERCHO.prop)

# Rename the columns for clarity
colnames(HYPERCHO.prop_df) <- c("Diagnosis", "HYPERCHO", "Proportion")

print(HYPERCHO.prop_df)

ggplot(HYPERCHO.prop_df, aes(x = HYPERCHO, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Hypercholesterolemia") +
  ylab("Diagnosis") +
  ggtitle("Hypercholesterolemia by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```
  
# Mosaic Plot of Race and Diagnosis
```{r}
# Calculate the proportion table
NACCNIHR_updated.prop <- prop.table(table(X18$NACCALZD, X18$NACCNIHR_updated), margin = 1)

# Convert the proportion table to a data frame
NACCNIHR_updated.prop_df <- as.data.frame(NACCNIHR_updated.prop)

# Rename the columns for clarity
colnames(NACCNIHR_updated.prop_df) <- c("Diagnosis", "NACCNIHR_updated", "Proportion")

print(NACCNIHR_updated.prop_df)

ggplot(NACCNIHR_updated.prop_df, aes(x = NACCNIHR_updated, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Race") +
  ylab("Diagnosis") +
  ggtitle("Race by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of Sex and Diagnosis
```{r}
# Calculate the proportion table
SEX.prop <- prop.table(table(X18$NACCALZD, X18$SEX), margin = 1)

# Convert the proportion table to a data frame
SEX.prop_df <- as.data.frame(SEX.prop)

# Rename the columns for clarity
colnames(SEX.prop_df) <- c("Diagnosis", "SEX", "Proportion")

print(SEX.prop_df)

ggplot(SEX.prop_df, aes(x = SEX, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Sex") +
  ylab("Diagnosis") +
  ggtitle("Sex by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```

# Mosaic Plot of Hispanic and Diagnosis
```{r}
# Calculate the proportion table
HISPANIC.prop <- prop.table(table(X18$NACCALZD, X18$HISPANIC), margin = 1)

# Convert the proportion table to a data frame
HISPANIC.prop_df <- as.data.frame(HISPANIC.prop)

# Rename the columns for clarity
colnames(HISPANIC.prop_df) <- c("Diagnosis", "HISPANIC", "Proportion")

print(HISPANIC.prop_df)

ggplot(HISPANIC.prop_df, aes(x = HISPANIC, y = Diagnosis, fill = Proportion)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Proportion * 100, 1), "%")),
            color = "black",
            size = 3,
            fontface = "bold",
            vjust = 0.5) +
  scale_fill_gradient(low = "#009E73", high = "#385a93", name = "Proportion") +
  xlab("Hispanic") +
  ylab("Diagnosis") +
  ggtitle("Hispanic by Diagnosis") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) + 
  theme(text = element_text(size = 14), panel.background = element_rect(fill = "#f5f0ec"), rect = element_rect(fill = "#f5f0ec", colour = "white",
                      size = 0.5, linetype = 1))
```